---
title: "test::curious"
subtitle: neet testing in research data science
author: Charles T. Gray </br> Interdisciplinary Metaresearch Group </br> School of History and Phisolophy of Science </br> University of Melbourne
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
  )
```

# *I'm* not checking your *code*! 

<div class="notes">
- How do you know I have?
- How do I know I have?
- Hubris to think our code is perfect
</div>

# test-driven research software engineering

<div class="notes">
We don't need to reinvent the wheel.
</div>

## research software engineering

<div class="notes">
Unit tests are one of many things that make up RSE
</div>

todo: get off the twitters

- **unit tests**
- containerisation

## research software engineering

<div class="notes">
Why are people adopting these tools?
</div>

### **containerisation**

todo: get Mark's thing off twitter

### **unit tests**

'I started using automated tests because I discovered I was spending too much time re-fixing bugs that I’d already fixed before.' - Testing, R Packages, Wickham 2015

## unit tests {.smaller}

<div class="notes">
Describe each of the three testing objects in relation to the example
</div>
 
```{r eval=FALSE, echo=TRUE}
context("String length")
library(stringr)

test_that("str_length is number of characters", {
  expect_equal(str_length("a"), 1)
  expect_equal(str_length("ab"), 2)
  expect_equal(str_length("abc"), 3)
})

test_that("str_length of factor is length of level", {
  expect_equal(str_length(factor("a")), 1)
  expect_equal(str_length(factor("ab")), 2)
  expect_equal(str_length(factor("abc")), 3)
})
```

testing object | description | function
- | - | -
**expectation** | atom of testing | `expect_*`
**test** | a collection of expectations | `test_that`
**context** | a collection of tests | `context`


# unit tests in data science | testing the good, the bad, and the ugly


<div class="notes">
Now we consider what we want to test in the context of data analysis
</div>


## testing the good | testing the good, the bad, and the ugly

<div class="notes">
Test for good inputs. Log expects a positive number.
</div>


```{r}
# test for fixed positive number
log(3)

# test for random positive number
x <- rexp(1)

# the exponential distribution only produces positive numbers

log(x)

```


## testing the bad| testing the good, the bad, and the ugly {.smaller}

<div class="notes">
Test for bad inputs. What if there's a numeric input, so expected type, but it's negative, so log doesn't work?
</div>


```{r}
# test for fixed negative number
log(-3)

# test for random netgative number
x <- -rexp(1)

# the exponential distribution only produces positive numbers

log(x)

```

## testing the ugly | testing the good, the bad, and the ugly

<div class="notes">
What about a non-numeric input, which is nonsense. We want it to fail & give informative errors.
</div>


```{r error = TRUE}
# test for ugly string input
log("pig")

# test for random string input
x <- c("cat", "pig", "bird") %>% sample(1)

log(x)

```

## test-driven development | balancing development with a measure of code::proof

todo: poster for the good the bad and the ugly

A **neet test** checks the function outputs a **non-empty** thing of **expected type**

# neet tests | a measure of sanity in building a data analysis

## neet test workflow

### set up

1. make a function
2. check in console it outputs a non-empty thing of expected type
3. convert this check into a test 
4. run test

```{r eval=FALSE}
usethis::use_testthat()


```


